---
title: "Time Series Analysis final Project - Competitive "
author: "Team Members: 1.Rahul k. gupta (sxxxxxxx) 2.Terrie christensen (sxxxxxxx) 3.Napapach Dechawatthanaphokin (s3613572)"
date: "14 May 2018"
output:
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: no
    toc_depth: 3
  html_document:
    df_print: paged
    toc: no
    toc_depth: '3'
  word_document:
    toc: no
    toc_depth: '3'
linkcolor: blue
references:
- author:
  - family: Xie
    given: YiHui
    issued:
    - year: 2018
    publisher: Chapman and Hall/CRC
    title: Dynamic Documents with R and knitr
  id: knitr
subtitle: MATH 1318 Time Series Analysis Final Project
documentclass: article
---

\newpage

\tableofcontents

\newpage

# Introduction \label{sec1}

Bitcoin or BTC is a digital currency, otherwise known as a cryptocurrency. it was introduced to markets in 2009 by 'Satoshi Nakamoto.' The most notable aspect of BTC is that no banks or financial institutions are needed to facilitate trades. Additionally, it runs as a virtually anonymous financial system whereby buyers and sellers do not need to input their names, addresses or any other personally identifiable information to transfer BTC.

Rahul Made these changes

# Model Diagnosis

```{r ,fig.width=8,fig.height=5,warning=FALSE,message=FALSE}
# Import Libraries
library(TSA)
library(fUnitRoots)
library(forecast)
library(CombMSC)
library(lmtest)
library(fGarch)
library(rugarch)

```


```{r ,fig.width=8,fig.height=5,warning=FALSE,message=FALSE,echo=FALSE}
# Require sort.score when knit to pdf, html
sort.score <- function(x, score = c("bic", "aic")){
  if (score == "aic"){
    x[with(x, order(AIC)),]
  } else if (score == "bic") {
    x[with(x, order(BIC)),]
  } else {
    warning('score = "x" only accepts valid arguments ("aic","bic")')
  }
}
```


```{r}
residual.analysis <- function(model, std = TRUE){
  library(TSA)
  library(FitAR)
  if (std == TRUE){
    res.model = rstandard(model)
  }else{
    res.model = residuals(model)
  }
  par(mfrow=c(3,2))
  plot(res.model,type='o',ylab='Standardised residuals', main="Time series plot of standardised residuals")
  abline(h=0)
  hist(res.model,main="Histogram of standardised residuals")
  qqnorm(res.model,main="QQ plot of standardised residuals")
  qqline(res.model, col = 2)
  acf(res.model,main="ACF of standardised residuals")
  print(shapiro.test(res.model))
  k=0
  LBQPlot(res.model, lag.max = length(model$residuals)-1 , StartLag = k + 1, k = 0, SquaredQ = FALSE)
  par(mfrow=c(1,1))
}
```


```{r}
Bitcoin <- read.csv("../Bitcoin_Historical_Price.csv", header=TRUE)
head(Bitcoin)

```



```{r}
Bitcoin <- ts(as.vector(Bitcoin$Close), frequency = 1)
class(Bitcoin)
Bitcoin.raw = Bitcoin
```


```{r}

plot(Bitcoin,type='o',ylab='Closing Price', main= "Time series plot for bitcoin")

```


```{r}
plot(y=Bitcoin,x=zlag(Bitcoin),ylab='Closing Price', xlab='Previous Year of bitcoin', main= "Scatter plot of bitcoin")
```


```{r}
y = Bitcoin             # Read the bitcoin data into y
x = zlag(Bitcoin)       # Generate first lag of the bitcoin data
index = 2:length(x)      # Create an index to get rid of the first NA value in x
cor(y[index],x[index])   # Calculate correlation between numerical values in x and y
```


```{r}
model.ln = lm(Bitcoin~time(Bitcoin)) # label the linear trend model as model.ln
summary(model.ln)
plot(ts(fitted(model.ln)), ylim = c(min(c(fitted(model.ln),
     as.vector(Bitcoin))), max(c(fitted(model.ln),as.vector(Bitcoin)))),
     ylab='y' , main = "Fitted quadratic curve to bitcoin data", type="l",lty=2,col="red")
lines(as.vector(Bitcoin),type="o")
```



```{r}
t = time(Bitcoin)
t2 = t^2
model.qa = lm(Bitcoin~ t + t2) # label the quadratic trend model as model.qa
summary(model.qa)
plot(ts(fitted(model.qa)), ylim = c(min(c(fitted(model.qa),
     as.vector(Bitcoin))), max(c(fitted(model.qa),as.vector(Bitcoin)))),
     ylab='y' , main = "Fitted quadratic curve to bitcoin data", type="l",lty=2,col="red")
lines(as.vector(Bitcoin),type="o")
```



```{r}
har.=harmonic(Bitcoin,0.5)
model.har=lm(Bitcoin~har.)
summary(model.har)
plot(ts(fitted(model.har)), ylim = c(min(c(fitted(model.har),
     as.vector(Bitcoin))), max(c(fitted(model.har),as.vector(Bitcoin)))),
     ylab='y' , main = "Fitted harmony curve to bitcoin data", type="l",lty=2,col="red")
lines(as.vector(Bitcoin),type="o")
```


# Models for Nonstationary Time Series

Auto regressive behaviour and non staionay
Staionay is the first thing we need to check.
```{r}
par(mfrow=c(1,2)) # To plot both ACF and PACF in the same panel of plots
acf(Bitcoin)
pacf(Bitcoin)
par(mfrow=c(1,1)) 
```


stategy to make stationay is transfromation.

```{r}
#Bitcoin.transform = BoxCox.ar(Bitcoin, method = "yule-walker") 
#Bitcoin.transform = BoxCox.ar(Bitcoin + abs(min(Bitcoin))+1) # You get an error saying the system is singular
#Bitcoin.transform = BoxCox.ar(Bitcoin + abs(min(Bitcoin))+1, lambda = seq(-1, 1, 0.01))
#Bitcoin.transform = BoxCox.ar(Bitcoin)
#Bitcoin.transform$ci
#lambda = 0.75 # 0.001
#Bitcoin.cash = (Bitcoin^lambda-1)/lambda #<--formula BoxCox transfromation
#plot(Bitcoin.cash,type='o',ylab='Los Angelos rainfall series')
```


```{r}
log.Bitcoin = log(Bitcoin)
plot(log.Bitcoin,type='o',ylab='Log-transformed of bitcoin data')
adf.test(log.Bitcoin)
```


```{r}
diff.log.Bitcoin = diff(log(Bitcoin), differences = 1)
#diff.log.Bitcoin = diff(log(Bitcoin))
plot(diff.log.Bitcoin,type='o',ylab='The first difference of log-transformed bitcoin')
adf.test(diff.log.Bitcoin)
```


```{r}
par(mfrow=c(1,2)) # To plot both ACF and PACF in the same panel of plots
acf(diff.log.Bitcoin, lag.max=50)
pacf(diff.log.Bitcoin, lag.max=50)
par(mfrow=c(1,1)) 
```


```{r}
eacf(diff.log.Bitcoin,ar.max = 12, ma.max = 12)
#ARIMA(0,1,0),ARIMA(0,1,1),ARIMA(0,1,2),ARIMA(0,1,3)
#ARIMA(1,1,1),ARIMA(0,1,2),ARIMA(0,1,3)
#ARIMA(2,1,2),ARIMA(0,1,3)
#ARIMA(3,1,3)
```

```{r}
res1 = armasubsets(y=diff.log.Bitcoin,nar=12,nma=12,y.name='test',ar.method='ols')
plot(res1)

```

```{r}

diff.log.Bitcoin2 = diff(log(Bitcoin), differences = 2)
par(mfrow=c(1,1))
plot(diff.log.Bitcoin2,type='o',ylab='Quarterly earnings ')
# There is still some trend in ther series
adf.test(diff.log.Bitcoin2)
```



```{r}
par(mfrow=c(1,2)) # To plot both ACF and PACF in the same panel of plots
acf(diff.log.Bitcoin2, lag.max=50)
pacf(diff.log.Bitcoin2, lag.max=50)
par(mfrow=c(1,1)) 
```


```{r}
eacf(diff.log.Bitcoin2,ar.max = 12, ma.max = 12)
#ARIMA(0,2,1),ARIMA(0,2,2),ARIMA(0,2,3),ARIMA(0,2,4)
#ARIMA(1,2,2),ARIMA(1,2,3),ARIMA(1,2,4)
#ARIMA(2,2,3),ARIMA(2,2,4),ARIMA(3,2,4)
```


```{r}
res2 = armasubsets(y=diff.log.Bitcoin2,nar=12,nma=12,y.name='test',ar.method='ols')
plot(res2)
```


```{r}
#The final set of possible models is 
#{#ARIMA(0,2,1),ARIMA(0,2,2),ARIMA(0,2,3),ARIMA(0,2,4)
#ARIMA(1,2,2),ARIMA(1,2,3),ARIMA(1,2,4)
#ARIMA(2,2,3),ARIMA(2,2,4),ARIMA(3,2,4)}

# ARIMA(0,2,1)
model_021_css = arima(log.Bitcoin,order=c(0,2,1),method='CSS')
coeftest(model_021_css)

model_021_ml = arima(log.Bitcoin,order=c(0,2,1),method='ML')
coeftest(model_021_ml)

# ARIMA(0,2,2)
model_022_css = arima(log.Bitcoin,order=c(0,2,2),method='CSS')
coeftest(model_022_css)

model_022_ml = arima(log.Bitcoin,order=c(0,2,2),method='ML')
coeftest(model_022_ml)

# ARIMA(0,2,3)
model_023_css = arima(log.Bitcoin,order=c(0,2,3),method='CSS')
coeftest(model_023_css)

model_023_ml = arima(log.Bitcoin,order=c(0,2,3),method='ML')
coeftest(model_023_ml)

# ARIMA(0,2,4)
model_024_css = arima(log.Bitcoin,order=c(0,2,4),method='CSS')
coeftest(model_024_css)

model_024_ml = arima(log.Bitcoin,order=c(0,2,4),method='ML')
coeftest(model_024_ml)

# ARIMA(1,2,2)
model_122_css = arima(log.Bitcoin,order=c(1,2,2),method='CSS')
coeftest(model_122_css)

model_122_ml = arima(log.Bitcoin,order=c(1,2,2),method='ML')
coeftest(model_122_ml)

# ARIMA(1,2,3)
model_123_css = arima(log.Bitcoin,order=c(1,2,3),method='CSS')
coeftest(model_123_css)

model_123_ml = arima(log.Bitcoin,order=c(1,2,3),method='ML')
coeftest(model_123_ml)

# ARIMA(1,2,4)
model_124_css = arima(log.Bitcoin,order=c(1,2,4),method='CSS')
coeftest(model_124_css)

model_124_ml = arima(log.Bitcoin,order=c(1,2,4),method='ML')
coeftest(model_124_ml)

# ARIMA(2,2,3)
model_223_css = arima(log.Bitcoin,order=c(2,2,3),method='CSS')
coeftest(model_223_css)

model_223_ml = arima(log.Bitcoin,order=c(2,2,3),method='ML')
coeftest(model_223_ml)

# ARIMA(2,2,4)
model_224_css = arima(log.Bitcoin,order=c(2,2,4),method='CSS')
coeftest(model_224_css)

model_224_ml = arima(log.Bitcoin,order=c(2,2,4),method='ML')
coeftest(model_224_ml)

# ARIMA(3,2,4)
model_324_css = arima(log.Bitcoin,order=c(3,2,4),method='CSS')
coeftest(model_324_css)

model_324_ml = arima(log.Bitcoin,order=c(3,2,4),method='ML')
coeftest(model_324_ml)

#{#ARIMA(0,2,1),ARIMA(0,2,2),ARIMA(0,2,3),ARIMA(0,2,4)
#ARIMA(1,2,2),ARIMA(1,2,3),ARIMA(1,2,4)
#ARIMA(2,2,3),ARIMA(2,2,4),ARIMA(3,2,4)}


```

```{r}
sort.score(AIC(model_021_ml,model_022_ml,model_023_ml,model_024_ml,model_122_ml,model_123_ml,model_124_ml,model_223_ml,model_224_ml,model_324_ml), score = "aic")
sort.score(BIC(model_021_ml,model_022_ml,model_023_ml,model_024_ml,model_122_ml,model_123_ml,model_124_ml,model_223_ml,model_224_ml,model_324_ml), score = "bic" )
```

```{r,fig.width=8,fig.height=5,warning=FALSE,message=FALSE,echo=FALSE}
residual.analysis(model = model_021_ml)
par(mfrow=c(1,1))
# There is no problem with the diagnostic plots.

# We applied log transformtion and second difference. To take them back:
log.data = log(Bitcoin.raw)
log.data.diff2.back = diffinv(diff.log.Bitcoin2, differences = 2, xi = data.matrix(log.data[1:2]))
log.data.diff2.back = exp(log.data.diff2.back)
```


```{r}
library(forecast)
fit = Arima(Bitcoin.raw,c(0,2,1)) 
plot(forecast(fit,h=10))
```

```{r}
McLeod.Li.test(y=diff.log.Bitcoin2,main="McLeod-Li Test Statistics for Daily Google Returns")
# McLeod-Li test is significnat at 5% level of significance for all lags. This gives a strong idea about existence of volatiliy clustering.
qqnorm(diff.log.Bitcoin2,main="Q-Q Normal Plot of Daily Google Returns")
qqline(diff.log.Bitcoin2) # Fat tails is in accordance with volatiliy clustering
```

```{r}
par(mfrow=c(1,2))
acf(diff.log.Bitcoin2, main="The sample ACF plot for return series")
pacf(diff.log.Bitcoin2, main="The sample PACF plot for return series")
eacf(diff.log.Bitcoin2)
```

```{r}
abs.diff.log.Bitcoin2 = abs(diff.log.Bitcoin2)
sq.diff.log.Bitcoin2 = diff.log.Bitcoin2^2
```


```{r}
par(mfrow=c(1,2))
acf(abs.diff.log.Bitcoin2, ci.type="ma",main="The sample ACF plot for absolute return series")
pacf(abs.diff.log.Bitcoin2, main="The sample PACF plot for absolute return series")
eacf(abs.diff.log.Bitcoin2)
```


```{r}
par(mfrow=c(1,2))
acf(sq.diff.log.Bitcoin2, ci.type="ma",main="The sample ACF plot for squared return series")
pacf(sq.diff.log.Bitcoin2, main="The sample PACF plot for squared return series")
eacf(sq.diff.log.Bitcoin2)
```

```{r}
m.11 = garch(diff.log.Bitcoin2,order=c(1,1),trace = FALSE)
summary(m.11) # All the coefficients are significant at 5% level of significance.
m.11_2 = garchFit(formula = ~garch(1,1), data =diff.log.Bitcoin2 )
summary(m.11_2)
```

